---
title: "Variable Selection"
output:
  html_document:
    theme: flatly
    toc: yes
    toc_float: yes
    font:family: Lato
  pdf_document:
    toc: yes
---

# Best Subset Selection

We apply the best subset selection to the `Hitters` data. We wish to predict a baseball player's `Salary` on the basis of various statistics associated with performance in the previous year.

First of all, we note that `Salary` variable is missing for some of the players. The `is.na()` function can be used to identify the missing observations. It returns a vector of the same length as the input vector, with a `TRUE` for any elements that are missing, and a `FALSE` for non-missing elements. The `sum()` function can then be used to count all of the missing elements.

```{r}
library(ISLR2)
```

Let's have a look at the variable names:

```{r}
names(Hitters)
```

We can also view the dataset dimensions:

```{r}
dim(Hitters)
```

Finally, let's identify any missing data:

```{r}
sum(is.na(Hitters$Salary))
```

We can see that `Salary` is missing for `59` players. The `na.omit()` function removes all of the rows that have missing values in any variable.

```{r}
hitters <- na.omit(Hitters)
sum(is.na(hitters))
```

We now have no `NA` values. 

The `regsubsets()` function (part of the `leaps` library) performs best subset selection by identifying the best model that contains a given number of predictors, where *best* is quantified using RSS. The syntax is the same as for `lm()`. The `summary()` command outputs the best set of variables for each model size.

```{r}
library(leaps)
regfit_full <- regsubsets(Salary ~ ., hitters)
summary(regfit_full)
```

An asterisk indicates that a given variable is included in the corresponding model. For instance, this output indicates that the best two-variable model contains only `Hits` and `CRBI`. By default, `regsubsets()` only reports results up to the best eight-variable model. But the `nvmax` argument can be used in order to return as many variables as are desired. Herem we fit up to a 19-variable model.

```{r}
regfit_full <- regsubsets(Salary ~ ., data = hitters, nvmax = 19)
regfit_sum <- summary(regfit_full)
```

The `summary()` function also returns *R2*, RSS, adjusted *R2*, *Cp*, and BIC. We can examine these to try and select the *best* overall model.

For instance, we can see that the *R2* statistic increases from 32% when only one variable in included in the model, to almost 55% when all variables are included. As expected, the *R2* statistic increases monotonically as more variables are included.

```{r}
regfit_sum$rsq
```

Plotting RSS, adjusted *R2*, *Cp*, and BIC for all of the models at once will help us decide which model to select. Note the `type = l` option tells `R` to connect the plotted points with lines.

```{r}
par(mfrow = c(1, 2))
plot(regfit_sum$rss, xlab = "Number of Variables",
    ylab = "RSS", type = "l")
plot(regfit_sum$adjr2, xlab = "Number of Variables",
    ylab = "Adjusted RSq", type = "l")
```

The `points()` command works like the `plot()` command, except that it puts points on a plot that has already been created, instead of creating a new plot. The `which.max()` function can be used to identify the location of the maximum point of a vector. We will now plot a red dot to indicate the model with the largest adjusted *R2* statistic

```{r}
which.max(regfit_sum$adjr2)
```

```{r}
plot(regfit_sum$adjr2, xlab = "Number of variables",
    ylab = "Adjusted Rsq", type = "l")
points(11, regfit_sum$adjr2[11], col = "red", cex = 2,
    pch = 20)
```

In a similar fashion, we can plot for Cp and BIC statistics, and indicate the models with the smallest statistic using `which.min()`.

```{r}
par(mfrow = c(2, 3))
plot(regfit_sum$rsq, xlab = "Number of variables",
    ylab = "Rsq", type = "l")
points(which.max(regfit_sum$rsq), regfit_sum$rsq[which.max(regfit_sum$rsq)],
    col = "red", cex = 2, pch = 20)
plot(regfit_sum$adjr2, xlab = "Number of variables",
    ylab = "Adjusted Rsq", type = "l")
points(which.max(regfit_sum$adjr2),
    regfit_sum$adjr2[which.max(regfit_sum$adjr2)],
    col = "red", cex = 2, pch = 20)
plot(regfit_sum$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l")
points(which.min(regfit_sum$cp), regfit_sum$cp[which.min(regfit_sum$cp)],
    col = "red", cex = 2, pch = 20)
plot(regfit_sum$rss, xlab = "Number of Variables",
    ylab = "Residual Sums Squared", type = "l")
points(which.min(regfit_sum$rss), regfit_sum$rss[which.min(regfit_sum$rss)],
    col = "red", cex = 2, pch = 20)
plot(regfit_sum$rss, xlab = "Number of Variables",
    ylab = "BIC", type = "l")
points(which.min(regfit_sum$bic), regfit_sum$bic[which.min(regfit_sum$bic)],
    col = "red", cex = 2, pch = 20)
```

The `regsubsets()` function has a built in `plot()` command which can be used to dispaly the selected variables for the best model with a given number of predictors.

```{r}
plot(regfit_full, scale = "r2")
```

The top row of each plot contains a black square for each variable selected according to the optimal model associated with that statistic. For instance, we see that several models share a BIC close to -150. Howeverm the model with the lowest BIC is the six-variable. We can use the `coef()` function to see the coefficient estimates associated with this model.

```{r}
coef(regfit_full, 6)
```

# Forwards and Backwards Stepwise Selection

We can also use the `regsubsets()` function to perform forward or backwise stepwise selection, using the argument `method = "forward` or `method = "backward"`.

```{r}
regfit_fwd <- regsubsets(Salary ~ ., data = hitters,
    nvmax = 19, method = "forward")
summary(regfit_fwd)
```

```{r}
regfit_bwd <- regsubsets(Salary ~ ., data = hitters,
    nvmax = 19, method = "backward")
summary(regfit_bwd)
```

For instance, we see that using forward stepwise selection tells us the best one-variable model contains only "CRBI", and the best two-variable model additionally includes `Hits`. The best seven-variable models identified by forward and backwards selection differ.

# Choosing Among Models Using the Validation-Set Approach and Cross-Validation

We previously saw that it is possible to choose among a set of models of different sizes using different criterions. We will now consider how to do this using the validation set and cross-validation approaches.

In order for these approaches to yield accurate estimates of the test error, we must use only the training observations to perform all aspects of model-fitting - including variable selection. Therefore, the determination of which model of a given size is best must be made using *only the training observations*. The point is subtle but important. If the full data set is used to perform the best subset selection step, the validation set errors and cross-validation errors that we obtain will not be accurate estimates of the test error.

In order to use the validation set approach, we begin by splitting the observations into a training set and a test set. We do this by creating a random vector, `train` of elementents equal to `TRUE` if the corresponding observation is in the training set, and `FALSE` otherwise. The vector `test` has a `TRUE` if the observation is in the test set, and a `FALSE` otherwise. Note the `!` in the command to create `test` causes the true and false to be switched around. We also set a random seed so that the user will obtain the same training set split.

```{r}
set.seed(42)
train <- sample(c(TRUE, FALSE), nrow(hitters), replace = TRUE)
test <- (!train)
```

Now, we apply `regsubsets()` to the training set in order to perform best subset selection.

```{r}
regfit_best <- regsubsets(Salary ~ ., data = hitters[train, ], nvmax = 19)
```

Notice that we subset the `hitters` data frame directly in the call in order to access only the training subset of the data, using the expression `hitters[train, ]`. We now compute the validation set error for the best model of each model size. We first make a model matrix from the test data.

```{r}
test_mat <- model.matrix(Salary ~ ., data = hitters[test, ])
```

The `model.matrix()` function is used in many regression packages for building an `X` matrix from data. Now, we run a loop, and for each size `i`, we extract the coefficients from `regfit_best` for the best model of that size, multiply them into the appropriate columns of the test model matrix to form the predictions, and compute the test Models

```{r}
val_err <- rep(NA, 19)
for (i in 1:19) {
    coefi <- coef(regfit_best, id = i)
    pred <- test_mat[, names(coefi)] %*% coefi
    val_err[i] <- mean((hitters$Salary[test] - pred)^2)
}
```

We find that the best model is the one that contains five variables.

```{r}
which.min(val_err)
```

```{r}
coef(regfit_best, 5)
```

This was a bit tedious, partly becuase there is no `predict()` method for `regsubsets()`. Since we will be using this function again, we can capture our steps above and write our own predict method.

```{r}
predict.regsubsets <- function(object, newdata, id, ...) {
    form <- as.formula(object$call[[2]])
    mat <- model.matrix(form, newdata)
    coefi <- coef(object, id = id)
    xvars <- names(coefi)
    mat[, xvars] %*% coefi
}
```

Our function pretty much mimics what we did above. The only complex part is how we extracted the formula used in the call to `regsubsets()`. We demonstrate how we use this function below, when we do cross-validation.

Finally, we perform best subset selection on the full data set, and select the best seven-variable model. It is important that we make use of the full data set in order to obtain more accurate coefficient estimates. Note that we perform best subset selection on the full data set and select the best five-variable model, rather than simply using the variables that were obtained from the training set, because the best five-variable model on the full data set may differ from the corresponding model on the training set. 

```{r}
regfit_best <- regsubsets(Salary ~ ., data = hitters, nvmax = 19)
coef(regfit_best, 5)
```

In fact, we see that the best seven-variable model on the full data set has a different set of variables than the best seven-variable model on the training set.

We now try to choose among the models of different sizes using cross-validation. This approach is somewhat involved, as we must perform best subset selection *within each of the K training sets*. Despite this, we see that with its clever subsetting syntax, `R` makes this job quite easy. First, we create a vector that allocates each observation to one of *k = 10* folds, and we create a matrix in which we store the results.

```{r}
k <- 10
n <- nrow(hitters)
set.seed(42)
folds <- sample(rep(1:k, length = n))
cv_errors <- matrix(NA, k, 19,
    dimnames = list(NULL, paste(1:19)))
```

Now, we write a loop that performs cross-validation. In the *j*th fold, the elements of `folds` that equal *`j`* are in the test set, and the remainder are in the training set. We make our predictions for each model suze (using our new `predict()` method), compute the test errors on the appropriate subset, and store them in the appropriate slot in the matrix `cv_errors`. Note that in the following code, `R` will automatically use our `predict.regsubsets()` function when we call `predict()` because the `best_fit` object has class `regsubsets`.

```{r}
for (j in 1:k) {
    best_fit <- regsubsets(Salary ~ ., data = hitters[folds != j, ],
        nvmax = 19)
    for (i in 1:19) {
        pred <- predict(best_fit, hitters[folds == j, ], id = i)
        cv_errors[j, i] <-
            mean((hitters$Salary[folds == j] - pred)^2)
    }
}
```

```{r}
dim(cv_errors)
```

This has given us a 10 x 19 matrix of which the *(j,i)* element corresponds to the test MSE for the *j*th cross-validation fold for the best *i*-variable model. We use the `apply()` function to average over the columns of this matrix in order to obtain a vector for which the *i*th element is the cross-validation error for the *i*-variable model.

```{r}
mean_cv_err <- apply(cv_errors, 2, mean)
mean_cv_err
```

```{r}
par(mfrow = c(1, 1))
plot(mean_cv_err, type = "b")
```

We see that cross-validation selects a 10-variable model. We now perform best subset selection on the full data set in order to obtain the 10-variable model.

```{r}
reg_best <- regsubsets(Salary ~ ., data = hitters, nvmax = 19)
coef(reg_best, 10)
```

## Subset Selection

```{r}
colnames(reg_data)
```

```{r}
# 0 Inputs
mod_0 <- lm(nox ~ 1, data = reg_data)

# 1 Input
age_fit <- lm(mpg ~ poly(weight, 4), data = Auto)
horse_fit <- lm(mpg ~ poly(horsepower, 2), data = Auto)
cylind_fit <- lm(mpg ~ cylinders, data = Auto)
which.min(c(deviance(weight_fit), deviance(horse_fit), deviance(cylind_fit)))

# poly(weight, 4) is the winner

# 2 Inputs

weight_horse_fit <- lm(mpg ~ poly(weight, 4) + poly(horsepower, 2), data = Auto)
weight_cyl_fit <- lm(mpg ~ poly(weight, 4) + cylinders, data = Auto)
horse_cyl_fit <- lm(mpg ~ poly(horsepower, 2) + cylinders, data = Auto)
which.min(c(deviance(weight_horse_fit), deviance(weight_cyl_fit), deviance(horse_cyl_fit)))

# poly(weight, 4) + poly(horsepower, 2) is the winner

# 3 Inputs

full_fit <- lm(mpg ~ poly(weight, 4) + poly(horsepower, 2) + cylinders, data = Auto)

# Cross Validation
Auto$const <- rep(0, dim(Auto)[1])

fit_1 <- train(mpg ~ const, data = Auto, method = "lm", trControl = train_control)
fit_2 <- train(mpg ~ poly(weight, 4), data = Auto, method = "lm", trControl = train_control)
fit_3 <- train(mpg ~ poly(weight, 4) + poly(horsepower, 2), data = Auto, method = "lm", trControl = train_control)
fit_4 <- train(mpg ~ poly(weight, 4) + poly(horsepower, 2) + cylinders, data = Auto, method = "lm", trControl = train_control)

mod_comp <- data.frame(mse = c(fit_1$results$RMSE, fit_2$results$RMSE, fit_3$results$RMSE, fit_4$results$RMSE),
                      Inputs = c(0, 1, 2, 3))

plot(mod_comp$Inputs, mod_comp$mse, type = "b", xlab = "Number of Input Variables", ylab = "RMSE")
```