```{r reprex-options, include = FALSE}
options(
  keep.source = TRUE,
  rlang_backtrace_on_error_report = "full",
  crayon.enabled = FALSE,
  reprex.current_venue = "gh"
)
```

```{r, results = 'asis', echo = FALSE, include = file.exists('.Rprofile'), eval = file.exists('.Rprofile')}
cat(sprintf("*Local `.Rprofile` detected at `%s`*", normalizePath(".Rprofile")))
```

---
output: reprex::reprex_document
knit: reprex::reprex_render
---

First, we load in and attach our our dataset

```{r}
library(ISLR2)
names(Smarket)
```

```{r}
dim(Smarket) # View the number of rows and columns
```

We can use the `pairs()` function to create a scatterplot matrix of all the variables.

The `cor()` function produces a matrix of all the pairwise correlations among the predictors.
We have to omit `direction` as it is a qualitative predictor

```{r}
pairs(Smarket)
cor(Smarket[, -9])
```

We can see that the correlations between `lag` variables and `Today` returns are close to zero,
In other words, there appears to be little correlation between today's returns and the previous days' returns.
The only substantial correlation is between `year` and `volume`.

<sup>Created on `r Sys.Date()` with [reprex v`r utils::packageVersion("reprex")`](https://reprex.tidyverse.org)</sup>
