---
output: reprex::reprex_document
knit: reprex::reprex_render
---

---
output: reprex::reprex_document
knit: reprex::reprex_render
---
# Libraries

Load in our packages

```{r}
library(MASS) # Contains our data
library(ISLR2) # Contains the Boston data set
```

Let's glimpse our dataset

```{r}
head(Boston)
```

Output variable:
- medv - medium house value. Our predictor variable

Input variables:
- rmvar - average number of rooms per house
- age average age of house 
- lstat - percent of households with low socioeconomic status

# Simple Linear Regression

Let's use the `attach` command to lock on our data set

```{r}
attach(Boston)
```

We can use the `lm` function to fit a simple linear regression model

```{r}
lm_fit <- lm(medv ~ lstat)
```

We can view some of the characteristics of the model
```{r}
lm_fit
```

For more detail information, we use `summary` function to give us the *p*-values and standard errors for the coefficients and the R^2 statistic and *F*-statistics

```{r}
summary(lm_fit)
```

We can use the `names()` function to find out more information in our model. 

```{r}
names(lm_fit)
```

We can also use the `coef()` function to find out the coefficients

```{r}
coef(lm_fit)
```

The `confint()` function shows us confidence intervals

```{r}
confint(lm_fit)
```

The `predict()` function shows us CI and prediction intervals for the prediction of `medv` for a given value of `lstat`

```{r}
predict(lm_fit, data.frame(lstat = (c(5, 10, 50))),
    interval = "confidence")
```

```{r}
predict(lm_fit, data.frame(lstat = (c(5, 10, 15))),
    interval = "prediction")
```

